---
title: "Class Project 1"
author: "Ed Yang"
date: "April 15, 2016"
output: html_document
---




## 1. Reading in the dataset and/or processing the data

```{r echo = TRUE}

    activity<-read.csv('./activity.csv')
    summary(activity)
    activity$dow <- weekdays(as.Date(activity$date))
    activity1 <- activity[!is.na(activity$steps),]
    library(plyr)
```

## 2. Histogram of the total number of steps taken each day



```{r echo = TRUE} 
    steptable <- aggregate(activity1$steps ~ activity1$date, FUN=sum )
    colnames(steptable)<- c("Date", "Steps")
    hist(steptable$Steps, breaks=8, xlab="Steps", main = "Total Steps per Day")
```

## 3. What is mean total number of steps taken per day?

```{r echo = TRUE}
    #Mean steps
    mean(steptable$Steps)
    #Median Steps
    median(steptable$Steps)
```    
## 4. What is the average daily activity pattern? 

```{r echo = TRUE}
    library(plyr)
    library(ggplot2)

    ##create average number of steps per interval
    intervals <- ddply(activity1, .(interval), summarize, Avg = mean(steps))
    
    ##Create line plot of average number of steps per interval
    p <- ggplot(intervals, aes(x=interval, y=Avg), xlab = "Interval", ylab="Average Number of Steps")
    p + geom_line()+xlab("Interval")+ylab("Average Number of Steps")+ggtitle("Average Number of Steps per Interval")
```



Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo = TRUE}
    maxst <- max(intervals$Avg)
    intervals[intervals$Avg==maxst,1]
```
The maximum number of steps for a 5-minute interval was 206.17 steps.

The 5-minute interval which had the maximum number of steps was the 835 interval.

## Imputing missing values

Calculate and report the total number of missing values in the dataset 

Devise a strategy for filling in all of the missing values in the dataset. 

Create a new dataset that is equal to the original dataset but with the missing data filled in.

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r echo=TRUE}
    #Number of NAs in original data set
    activity2<-activity[is.na(activity$steps),]
    summary(activity2)
```
The length of the new table is 2304, so we have 2304 missing data records

```{r echo=TRUE}
    library(magrittr) 
    library(dplyr)

    activityt<- activity
    nas<- is.na(activityt$steps)
    avg_interval<- tapply(activityt$steps, activityt$interval, mean, na.rm=TRUE, simplify = TRUE)
    activityt$steps[nas] <- avg_interval[as.character(activityt$interval[nas])]

    Total_Steps2<- activityt%>%
    group_by(date)%>%
    summarise(total_steps2 = sum(steps, na.rm=TRUE))

    hist(as.numeric(Total_Steps2$total_steps), breaks=8, xlab="Steps", main = "Total Steps per Day")
```

## Are there differences in activity patterns between weekdays and weekends

```{r echo=TRUE}
    
    library(lattice)
    activityt$dow1 <- ifelse(activityt$dow %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
    
    finaltable <- ddply(activityt, .(interval, dow1), summarize, Avg = mean(steps))

    xyplot(Avg~interval|dow1, data=finaltable, type="l",  layout = c(1,2),
      main="Average Steps per Interval Based on Type of Day", 
     ylab="Average Number of Steps", xlab="Interval")

```
Yes, people do walk more in weekends!!

    